---
title: "Create semantic expressions"
author: "Jasper Ginn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Cortical API requires semantic expression to be written in a format that is not native to R. This document shows the general procedure to construct semantic expressions in R. For more information, check the [Cortical documentation](http://documentation.cortical.io/the_power_of_expressions.html)

## Constructing semantic expressions

Take a look at the following two examples from the Cortical documentation:

```json
{
     "term" : "apple"
}
```

and

```json
{
  "sub" : [
    {
      "term" : "apple"
    },
    {
      "text" : "Mac OS is a series of graphical user interface-based operating systems developed by Apple Inc. for their Macintosh line of computer systems."
    }
  ]
}
```

This looks (and is) a bit complicated, but in general, the rules are like this:

1. If you use one term, text or array of positions, then you can use a simple named list:

```{r}
exp <- list(
  "term" = "apple"
)
```

2. If you use an operator, you should name the operator, its value should be a list, and each element of this list should be a list:

```{r}
exp <- list(
  # Named operator
  "sub" = list( # ... whose value is a list
    list("term" = "apple"), # ... each element has its own list.
    list("text" = "Mac OS is a series of graphical user interface-based operating systems developed by Apple Inc. for their Macintosh line of computer systems.")
  )
)
```

3. If you want to POST multiple expressions to the API, then you must construct a list in which each individual expression is an unnamed list:

```{r}
exp <- list(
  # Each element is an unname list
  list(
    "term" = "apple"
  ),
  # Second element
  list(
    "sub" = list( 
      list("term" = "apple"), 
      list("text" = "Mac OS is a series of graphical user interface-based operating systems developed by Apple Inc. for their Macintosh line of computer systems.")
    )
  )
)
```
