---
title: "Create semantic expressions"
author: "Jasper Ginn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Cortical API requires semantic expression to be written in a format that is not native to R. This can be a problem if you need to create a slightly more complex expression. `FinTxtUtils` contains a couple of useful functions to help you construct semantic expressions.

For the full explanation, execute `?FinTxtUtils::create_expression` in R.

To begin building an expression, you call `create_expression()`, which you then populate using e.g. `add_term()` and operators such as `AND()`, `OR()` etc. If you call an operator, you need to 'close' it using the `END()` operator.

```{r}
library(magrittr)
library(FinTxtUtils)

# Start building the expression
exp <- create_expression() %>%
  # Subtract the following
  SUB() %>%
  add_term("apple") %>%
  AND() %>%
  # The following entries are added to the "SUB" level until you call "END()"
  add_text("Mac OS is a series of graphical user interface-based operating systems developed by Apple Inc. for their Macintosh line of computer systems.") %>%
  add_term("banana") %>%
  END() %>%
  END()

knitr::kable(exp@expression)
```

You can build the expression using the `build_expression()` command

```{r}
exp %>%
  build_expression()
```

As you can see, this is a messy query, but `do_fingerprint_expression()` knows how to handle it:

```{r, echo=FALSE}
# THIS IS TEMPORARY
io <- exp %>% build_expression()
io <- io[[1]]
```

```{r}
expr <- do_fingerprint_expression(io)
```
