---
title: "Use FactoMineR for dimension reduction"
author: "Jasper Ginn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Because the retina returns binary values in a finite space (16384 positions), doing computations is more efficient than many other methods. However, the more documents you have, the more expensive computation becomes, especially if you want to compare documents to one another.

Fortunately, we can reduce the number of dimensions of the retina dramatically while retaining sensible results. This vignette shows how this can be done. 

```{r}
suppressPackageStartupMessages(library(sfutils))
# Load data
data("SAP500")
```

We will make a subset of this data (50 companies) for the purposes of this document:

```{r}
# Choose these companies
chosen <- c(1:3, 12:13,
            16, 21, 28:29, 
            33, 35, 60:62, 
            69, 80, 119, 121,
            149, 160, 168, 174,
            183, 186, 193, 205,
            219, 229, 235, 262,
            266, 278, 279, 338,
            339, 342, 343, 353, 
            355, 359, 362, 373,
            378, 387, 409, 435, 
            436, 453, 468, 487) # 366
# Subset
subs <- SAP500[chosen]
names(subs)
```

We fingerprint the descriptions, calculate a distance matrix and cluster the results.

```{r, eval=FALSE}
# Retrieve only descriptions
desc_only <- lapply(subs, function(x) x$desc)

# Fingerprint
fps <- do_fingerprint_document(desc_only)

# To matrix
binmat <- as.matrix(fps)
rownames(binmat) <- names(desc_only)

# Distances (need to convert matrix from Matrix package to normal matrix)
fpsdst <- dist(as.matrix(binmat))

# Do hierarchical clustering and plot
clust <- hclust(fpsdst, method = "ward.D2")
```

<img src=dendro1.png width=80%/>

We now do the same but using PCA to reduce the number of columns to 30

```{r, eval = FALSE}
# Reduce dimensions
library(FactoMineR)
red <- PCA(as.matrix(binmat),
           ncp = ncol(binmat), graph = FALSE)

# Predict
pred <- predict(red, binmat)$coord

# Take 64 dims
pred_filt <- pred[,1:30]

# As distance
dd <- dist(pred_filt)

# Cluster
clusts <- hclust(dd, method = "ward.D2")
```

<img src=dendro2.png width=80%/>
